---
title: "Iris species identification"
author: "Julian Störig"
date: "17.08.2020"
output: html_document
---

## Data Description

The iris data-set contains 150 observation, divided into
50 observations each between three different species of iris flower,
Iris setosa, Iris versicolor and Iris virginica.

For each observation it includes the length and width of the flowers
sepal and petal leaves, along with the species name.

## Data Analysis

The data analysis for this report is done in R, this report written in reproducible R-Markdown, the source code for which is availiable [here](https://github.com/julianstoerig/data-analysis-and-templates/blob/master/learning/IrisSpeciesIdentificationReport.Rmd).


```{r, echo=FALSE}
library(tibble)
library(ggplot2)
library(utils)
library(dplyr)

data <- as_tibble(iris)
```

### scatter plots showing all indifidual data points

To get a general feel for the data, it is often advisable to take a look at all datapoints, find outliers and try to recognize trends, etc..

#### sepal leaves by species

```{r, echo=FALSE}
ss <- ggplot(data, aes(x=Sepal.Length,
                 y=Sepal.Width,
                 color=Species)) +
  geom_point()
ss <- ss +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 2",
       subtitle = "petal leaves: scatter plot")
ss
```



#### petal leaves by species

```{r, echo=FALSE}
ps <- ggplot(data, aes(x=Petal.Length,
                 y=Petal.Width,
                 color=Species)) +
  geom_point()
ps <- ps +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 2",
       subtitle = "petal leaves: scatter plot")
ps
```

### box plots showing general trend for the different species

#### sepal leaves by species

```{r, echo=FALSE}
sb <- ggplot(data, aes(x=Sepal.Length,
                 y=Sepal.Width,
                 color=Species)) +
  geom_boxplot()
sb <- sb +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 3",
       subtitle = "sepal leaves: box plot")
sb
```



#### petal leaves by species

```{r, echo=FALSE}
pb <- ggplot(data, aes(x=Petal.Length,
                 y=Petal.Width,
                 color=Species)) +
  geom_boxplot()
pb <- pb +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 4",
       subtitle = "petal leaves: box plot")
pb
```

### conclusions

For the identification of these Iris species, the petal leaves are of much
higher significance than the sepal leaves. This is because the overlap between
I. versicolor and I. virginica is much larger for the sepal leaves (*fig. 1*),
than it is for the petal leaves (*fig. 2*).

Differentiating between these two and I. setosa is also easiest by
measuring the petal leaves, as is evidenced by the seperation between
I. setosa and the to other Iris species.
visible in *fig. 4*.