---
title: "Iris Species Identification"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tibble)
library(ggplot2)
library(utils)
library(dplyr)

data <- as_tibble(iris)
```


## Column {data-width=500, .tabset}

### Abstract

An Analysis of the Iris data-set reveals that the importance of the petal size
is much higher than that of the sepal leaf size for identification of these Iris
species, and also shows that to gain a higher identification accuracy, it is also
useful to take measurements of the sepal leaf width, which is much more useful
than sepal leaf length.

### Introduction

The iris data-set contains 150 observation, divided into
50 observations each between three different species of iris flower,
Iris setosa, Iris versicolor and Iris virginica.

For each observation it includes the length and width of the flowers
sepal and petal leaves, along with the species name.

The data analysis for this report is done in R, this report written in reproducible R-Markdown, the source code for which is availiable [here](https://github.com/julianstoerig/data-analysis-and-templates/blob/master/learning/IrisSpeciesIdentificationReport.Rmd).

### Data Analysis

For the identification of these Iris species, the petal leaves are of much
higher significance than the sepal leaves. This is because the overlap between
I. versicolor and I. virginica is much larger for the sepal leaves (*fig. 1*),
than it is for the petal leaves (*fig. 2*).

Differentiating between these two and I. setosa is also easiest by
measuring the petal leaves, as is evidenced by the seperation between
I. setosa and the to other Iris species.
visible in *fig. 4*.

## Column {data-width=500, .tabset}

### Sepal leaves: *fig. 1 & fig. 3*

```{r, echo=FALSE}
# figure 1
ss <- ggplot(data, aes(x=Sepal.Length,
                 y=Sepal.Width,
                 color=Species)) +
  geom_point()
ss <- ss +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 2",
       subtitle = "petal leaves: scatter plot")
ss

# figure 3
sb <- ggplot(data, aes(x=Sepal.Length,
                 y=Sepal.Width,
                 color=Species)) +
  geom_boxplot()
sb <- sb +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 3",
       subtitle = "sepal leaves: box plot")
sb
```

### Petal leaves: *fig. 2 & fig. 4*

```{r, echo=FALSE}
# figure 2
ps <- ggplot(data, aes(x=Petal.Length,
                 y=Petal.Width,
                 color=Species)) +
  geom_point()
ps <- ps +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 2",
       subtitle = "petal leaves: scatter plot")
ps

# figure 4
pb <- ggplot(data, aes(x=Petal.Length,
                 y=Petal.Width,
                 color=Species)) +
  geom_boxplot()
pb <- pb +
  xlab("length") +
  ylab("width") +
  labs(title = "fig. 4",
       subtitle = "petal leaves: box plot")
pb
```

